### 2018-10-26 _analyze.Rmd
Add data analyses and simulations

---
title: '`r paste("Experiment", params$experiment)`'
description: |
  Data analysis for Experiment `r params$experiment`.
author:
  - name: Matti Vuorre 
    affiliation: Columbia University
date: "`r Sys.Date()`"
output: 
  radix::radix_article:
    toc: true
    toc_depth: 1
    self_contained: false
    lib_dir: "docs/site_libs"
params:
  experiment: 3
---

```{r setup, include=FALSE}
source("functions.R")
library(knitr)
library(summarytools)
library(tidyverse)
library(broom)
library(scales)
library(patchwork)
library(lme4)
library(afex)
library(emmeans)
opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  eval = TRUE,
  dev = c("png", "pdf"),
  fig.ext = c("png", "pdf"),
  fig.retina = 2
)
theme_set(theme_linedraw() + theme(panel.grid = element_blank()))
```

# Data description

```{r create-data}
dat <- read_csv("data/merged.csv", guess_max = 25000)

# Get data for this experiment only
dat <- filter(dat, exp == params$experiment)

# Drop excluded subjects
dat <- filter(dat, !exclude)

# Drop factor levels of subject used in other experiments
dat$subject <- fct_drop(dat$subject)
```

```{r metrics-summarise, results = 'asis'}
dat_metrics <- dat %>%
  nest(-subject) %>%
  mutate(res = map(data, metrics, ordinal = TRUE)) %>%
  unnest(res, .drop = TRUE)
# Summarize
dat_metrics %>% 
  select(-contains("s")) %>%
  descr(stats = c("mean", "sd", "min", "q1", "q3", "max", "n.valid", "pct.valid"), 
        round.digits = 2, style = "rmarkdown")
```

# Math performance

ANOVA of proportion correct on session. We use same exclusion rules as above but at the session level. ANOVA only considers subjects who had observations at each session; thus we also fit a random intercept linear regression to include everyone.

```{r math-performance-sessions}
exclude_threshold <- 3

# Get gamma, confidence, and performance for everyone & sessions
dat_sessions <- dat %>%
  group_by(session, subject) %>%
  nest() %>%
  mutate(res = map(data, metrics)) %>%
  unnest(res)

# Create indicator for each of five exclusions rules for each subject. Variables are z-scored WITHOUT considering subjects with negative or NA gammas.
dat_sessions <- dat_sessions %>%
  mutate(
    e1 = is.na(gamma),
    e2 = gamma < 0,
    # Negative gammas to NAs so they don't impact next calculations
    gamma = ifelse(gamma < 0, NA, gamma),
    e3 = abs(scale(gamma)) > exclude_threshold,
    e4 = abs(scale(p)) > exclude_threshold,
    e5 = abs(scale(con_m)) > exclude_threshold
  ) %>%
  group_by(subject, session) %>%
  mutate(exclude = any(e1, e2, e3, e4, e5)) %>%
  ungroup() %>%
  filter(!exclude)
```

```{r math-performance-sessions-anova}
fit_aov_math_sessions <- aov_ez(
  id = "subject",
  dv = "p",
  data = dat_sessions,
  within = "session",
)
list(tidy(fit_aov_math_sessions$anova_table),
     tidy(fit_aov_math_sessions$aov)) %>% 
  kable(digits = 3, caption = "ANOVA tables (corrected & not corrected)")
paste("Subjects in ANOVA:", nrow(fit_aov_math_sessions$data$wide))
summary(lmer(p ~ session + (session || subject),
             data = dat_sessions))["coefficients"] %>%
  kable(digits = 3, caption = "LMER Table")
dat_sessions %>%
  ggplot(aes(session, p)) +
  geom_point(shape = 1, position = position_jitter(h = 0, w = .1)) +
  stat_smooth(method = "lm") +
  stat_summary(fun.data = mean_cl_boot)
```

# Resolution

## Gamma by session

We used the same rejection rules for gammas at the session level.

```{r gamma-sessions}
fit_aov_gamma_sessions <- aov_ez(
  id = "subject",
  dv = "gamma",
  data = dat_sessions,
  within = "session"
)
list(tidy(fit_aov_gamma_sessions$anova_table),
     tidy(fit_aov_gamma_sessions$aov)) %>% 
  kable(digits = 3, caption = "ANOVA tables (corrected & not corrected)")
paste("Subjects in ANOVA:", nrow(fit_aov_gamma_sessions$data$wide))
summary(lmer(gamma ~ session + (session || subject),
  data = dat_sessions))["coefficients"] %>%
  kable(digits = 3, caption = "Gamma LMER Table")
dat_sessions %>%
  ggplot(aes(session, gamma)) +
  geom_point(shape = 1, position = position_jitter(h = 0, w = .1)) +
  stat_smooth(method = "lm") +
  stat_summary(fun.data = mean_cl_boot)
```

## Overall gamma

```{r gamma-overall}
dat_metrics %>%
  summarise(
    mean = mean(gamma, na.rm = T),
    N = n(),
    se = sd(gamma, na.rm = T) / sqrt(N)
  ) %>%
  kable(digits = 2)
```

```{r gamma-subject-ttest}
t.test(dat_metrics$gamma)
```

## Type 2 SDT

In calculating da, we added .5 to all cells. We calculate $d_a$ by obtaining intercept and slope of linear regression of z-scores in right panel above. We also obtained SDT metrics via ordinal regression to confirm.

```{r da-ttest}
t.test(dat_metrics$da)
```

```{r resolution-scatterplots, fig.cap = "Average and 95pCIs gammas and das. Points = subjects."}
dat_metrics %>%
  select(subject, gamma, da) %>%
  gather(metric, value, gamma, da) %>%
  ggplot(aes(x = metric, y = value)) +
  scale_color_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, lty = 2, size = .33) +
  geom_point(
    shape = 21,
    fill = "white",
    alpha = .5,
    size = .75,
    position = position_jitter(h = 0, w = .075)
  ) +
  stat_summary(
    fun.y = mean,
    geom = "point",
    size = 2.5
  ) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = .33, size = .75
  ) +
  scale_y_continuous(breaks = pretty_breaks()) +
  facet_wrap("metric", scales = "free") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
```

# Resolution - performance Correlation

```{r gamma-performance-plot}
p_gamma_p <- dat_metrics %>%
  ggplot(aes(p, gamma)) +
  geom_vline(xintercept = .25, col = NA) + # Ensure limits
  geom_vline(xintercept = 1, col = NA) + # Ensure limits
  geom_hline(yintercept = 0, lty = 2, size = .33) +
  geom_point(shape = 21, fill = "white", alpha = .75) +
  geom_smooth(method = lm, col = "black", size = .75, alpha = .33) +
  scale_x_continuous(breaks = c(.25, .5, .75, 1)) +
  labs(
    x = "Proportion correct",
    y = "Gamma"
  )
p_gamma_p
```

We computed da using two different methods (as above and ordinal regression); displayed below:

```{r da-performance-plot}
dat_metrics %>% 
  gather(method, value, da, da_ord) %>% 
  ggplot(aes(p, value)) +
  geom_hline(yintercept = 0, lty = 2, size = .33) +
  geom_text(aes(label = subject)) +
  geom_smooth(method = lm, col = "black", size = .75, alpha = .33) +
  scale_x_continuous(breaks = c(.25, .5, .75, 1)) +
  scale_y_continuous(limits = c(0, 2)) +
  facet_wrap("method") +
  labs(
    x = "Proportion correct",
    y = expression(italic(d)[a])
  )

dat_metrics %>%
  ggplot(aes(da, da_ord)) +
  geom_point() +
  geom_abline() +
  theme(aspect.ratio = 1) |
  dat_metrics %>%
    ggplot(aes(y0, d_ord)) +
    geom_point() +
    geom_abline() +
    theme(aspect.ratio = 1)
```

Pearson correlation between gamma and performance

```{r gamma-performance-test}
cor.test(dat_metrics$gamma, dat_metrics$p, method = "pearson")
```

Between da and performance

```{r da-performance-test}
cor.test(dat_metrics$da, dat_metrics$p, method = "pearson")
```

# Calibration

## Overconfidence

Calculate bias scores.

```{r bias-calculate}
dat_bias <- dat %>%
  # Rescale confidence to 0-1 scale
  mutate(con_01 = ifelse(exp != 5, con6 / 5, con / 10)) %>%
  group_by(exp, subject) %>%
  summarise(
    p = mean(accuracy, na.rm = T),
    con_01 = mean(con_01, na.rm = T)
  ) %>%
  mutate(bias = con_01 - p)
```

Summarise average bias scores and test against zero:

```{r bias-summary-test}
dat_bias %>%
  group_by(exp) %>%
  summarise(
    mean = mean(bias, na.rm = T),
    n = n(),
    se = sd(bias, na.rm = T) / sqrt(n),
    NAs = sum(is.na(bias))
  ) %>%
  kable(digits = 3)
t.test(dat_bias$bias)
```

Univariate figure

```{r bias-plot, eval = F}
dat_bias %>%
  ggplot(aes(x = "bias", y = bias)) +
  scale_color_brewer(palette = "Set1") +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_point(
    shape = 1,
    alpha = .5,
    size = .66,
    position = position_jitter(h = 0, w = .05)
  ) +
  stat_summary(
    fun.y = mean,
    geom = "point",
    size = 2.5
  ) +
  stat_summary(
    fun.data = mean_cl_boot,
    geom = "errorbar",
    width = .33, size = .75
  ) +
  labs(y = "Bias") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none"
  )
```

## Bias - performance correlation

```{r bias-plot-performance}
p_bias_p <- dat_bias %>%
  ggplot(aes(p, bias)) +
  geom_vline(xintercept = .25, col = NA) + # Ensure limits
  geom_vline(xintercept = 1, col = NA) + # Ensure limits
  geom_hline(yintercept = 0, lty = 2, size = .33) +
  geom_point(shape = 21, alpha = .75, fill = "white") +
  geom_smooth(method = lm, col = "black", size = .75, alpha = .33) +
  scale_x_continuous(breaks = c(.25, .5, .75, 1)) +
  labs(
    x = "Proportion correct",
    y = "Bias"
  )
p_bias_p
```

```{r bias-performance-correlate}
# fit_bias <- lm(bias ~ scale(p), data = dat_bias)
# summary(fit_bias)
cor.test(dat_bias$bias, dat_bias$p)
```

# Publication figure

```{r figure-pub, fig.height = 3.4, fig.width = 6.6, cache = F}
(p_gamma_p + p_bias_p) *
  theme(aspect.ratio = 1) +
  plot_annotation(tag_levels = "A")
```