### 2019-05-08 index.Rmd

---
title: "Metacognition and mathematics performance"
date: "`r Sys.Date()`"
author:
  - name: Matti Vuorre 
    affiliation: Columbia University
output: 
  radix::radix_article:
    toc: true
    toc_depth: 1
    self_contained: false
  github_document:
    html_preview: false
site: radix::radix_website
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = FALSE)
```

Data and analyses related to Vuorre & Metcalfe (*"Metacognition about Mathematics is Excellent but the Relation between Performance and Resolution is Confounded by Guessing"*; in prep).

* Data from all six experiments are in a single table at `data/merged.csv`
* Analyses are documented as an [R Markdown](https://rmarkdown.rstudio.com/) ([Radix](https://rstudio.github.io/radix/)) website at `docs/index.html`, which can be viewed at <https://mvuorre.github.io/mathPerfConf>
* The materials are archived on OSF at <https://osf.io/np3cu/>, and github at <https://github.com/mvuorre/mathPerfConf>

## Reproduce

To reproduce these analyses, copy all files to a local drive. Activate the `mathPerfConf` R Project in R Studio, and run `build.R`. The analysis code is contained in `.Rmd` R Markdown files, which can also be executed individually in interactive mode.

## Data description

```{r, echo = TRUE, message = FALSE, collapse = TRUE}
dat <- read_csv("data/merged.csv", guess_max = 25000)
glimpse(dat, 1)
```

We binned confidence ratings to six bins (Experiments 1-3 already had six-point ratings, 4 & 6 were slider ratings, 5 had a 0-10 Likert scale). The raw confidence ratings are variable `con`, binned ratings are variable `con6`.

```{r, eval = F}
# Export list of items to table
dat %>% 
  group_by(exp, item) %>% 
  count() %>% 
  spread(exp, n, sep = "", fill = "-") %>% 
  write_csv("tab.txt")

dat %>% 
  distinct(exp, item) %>% 
  group_by(exp) %>% 
  mutate(problem = 1:n()) %>% 
  spread(exp, item, fill = " ") %>% 
  write_csv("tab.txt")
```