### 2019-05-08 simulation.Rmd

---
title: "Simulation"
description: |
  A simulation study investigating the impact of task knowledge and guessing on resolution metrics
author:
  - name: Matti Vuorre 
    affiliation: Columbia University
date: "`r Sys.Date()`"
output: 
  radix::radix_article:
    toc: true
    toc_depth: 2
    self_contained: false
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
source("functions.R")
library(ordinal)
library(knitr)
library(beepr)
library(parallel)
library(scales)
library(patchwork)
library(tidyverse)
library(furrr)
library(broom)
opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  error = TRUE,
  cache = FALSE,
  dev = c("png", "pdf"),
  fig.ext = c("png", "pdf"),
  fig.retina = 2
  )
theme_set(theme_linedraw() + 
            theme(panel.grid = element_blank(),
                  legend.position = "right"))
options(mc.cores = parallel::detectCores())
```

```{r data-load}
dat <- read_csv("data/merged.csv", guess_max = 25000)
# Drop excluded subjects
# dat <- filter(dat_full, !exclude)
# Drop unused factor levels
dat$subject <- fct_drop(dat$subject)

dat_metrics <- dat %>% 
  nest(-exp, -subject) %>% 
  mutate(res = map(data, metrics, ordinal=TRUE)) %>% 
  unnest(res, .drop = TRUE) %>% 
  # Abbott's formula to correct performance for guessing (Lawless p.11)
  mutate(p_know = (p - .25) / (1 - .25)) %>% 
  filter(p_know > 0)
```

## Preface

This simulation examines potential bias in measures of metacognitive resolution due to guessing and task knowledge.

It was previously argued that gamma is unrelated to criterion test performance @Nelsoncomparisoncurrentmeasures1984:

>Property 3. The feeling-of-knowing accuracy score should be independent of overall recognition performance.  
Property 3 states that there should not be any built-in relation between feeling-of-knowing accuracy and overall recognition. The measure of feeling-of-knowing accuracy should allow for the possibility of a person with perfect feeling-of-knowing accuracy at each of many levels of overall recognition performance; for example, there could be completely accurate feeling-of-knowing predictions for a person with overall recognition performance of 90% correct, completely accurate feeling-of-knowing predictions for a person with overall recognition performance of 80% correct, and so on.

However, in mAFC tasks, gamma is downward biased as a function of m (number of response alternatives) [@schwartz1994methodological]. Below, we show in detail how guesses cause this bias.

### Example

Consider two people with perfect resolution answering 10 4-alternative MC questions. Person A knows 2 items, and B knows six items. However, due to guessing, A gets 4/10 right (2 correct guesses) and B gets 7/10 right (1 correct guess). But their resolution is perfect. Metrics of resolution therefore should indicate so.

Here is example person A's table, observed gamma, and distribution of gammas over 1000 replications.

```{r example-a}
example <- function(know=2, guesses=2, levels=6, n=10) {
  # Must sample letters because sample(x) automatically draws numbers <= x
  lls <- letters[1:levels]
  tibble(
    Know = c(rep(0, n-know), rep(1, know)),
    Accuracy = c(rep(0, (n-know-guesses)), rep(1, guesses+know)),
    Confidence = ordered(
      ifelse(
        Know==0, 
        sample(lls[1:(levels/2)], n-know, T), 
        sample(lls[(levels/2+1):levels], know, T)
      ), 
      levels = lls, labels = 1:levels)
  )[,-1]
}
set.seed(555)
a <- example(2, 2, 6, 10)
table(a)
vcdExtra::GKgamma(table(a))$gamma
summary(replicate(1000, vcdExtra::GKgamma(table(example(2, 2, 6, 10)))$gamma))
```

And the same for B

```{r example-b}
b <- example(6, 1, 6, 10)
table(b)
vcdExtra::GKgamma(table(b))$gamma
summary(replicate(1000, vcdExtra::GKgamma(table(example(6, 1, 6, 10)))$gamma))
```

## Simulation overview

This simulation produces hypothetical data sets of responses (accuracy & confidence rating pairs) under various conditions of guessing, performance, and true resolution. The simulation scheme is adapted from @benjamin2008measurement, and is outlined for one hypothetical subject as follows:

- A list of items is generated based on number of items, and proportion of known items (Knowledge = *K*)
- For each item, accuracy is determined such that known items are always correct, whereas unknown items are correct according to a defined guessing probability *G*
- For each item, a "metacognitive evidence" value is generated from one of two gaussian distributions depending on whether the item was known or unknown
  - Unknown item evidence values: mean zero and standard deviation one
  - Known item evidence values: mean *R* and standard deviation $\sigma_k$. Thus, R defines the generative resolution parameter as the difference between the evidence distributions.
- Evidence values are converted to confidence ratings (6-category scale) values via five thresholds
- Resolution metrics are obtained based on this simulated data set

This simulation is then repeated over a set of hypothetical subjects to obtain mean estimated metrics for combinations of user-determined values. A summary of these variables appears in Table \@ref(tab:symbol-table).

```{r symbol-table}
tibble(
  Variable = c(
    "Number of trials/items", 
    "Proportion of known items (knowledge)",
    "Guessing proportion (i.e. 1/m)",
    "Mean of evidence distribution for known items",
    "SD of evidence distribution for known items",
    "Evidence thresholds"
  ),
  Symbol = c(
    "$N_{trials}$",
    "$K$",
    "$G$",
    "$R$",
    "$\\sigma_k$",
    "$\\tau$"
  ),
  `Default value` = c(
    100, 0.5, 0.0, 1.0, 1.0, "-1, -.5, 0, .5, 1"
  )) %>% 
  kable(
    escape = FALSE, 
    caption = "Summary of generative variables used in simulation function."
  )
```

Metacognitive resolution, defined as *R*, is then estimated with a number of metrics:

- Goodman-Kruskal Gamma correlation
- Signal-detection metrics $y_0$ (y-intercept of the type-2 zROC) and $d_a$ [@benjamin2008measurement]
- Signal-detection metrics equivalent to above, but obtained from an ordinal regression [@BurknerOrdinalRegressionModels2018; @KnoblauchModelingPsychophysicalData2012], $d_{a-ord}$ and $d'_{ord}$
- Pearson correlation $r$

<aside>
$d_a = \frac{\sqrt{2}y_0}{\sqrt{1 + m^2}}$, where $m$ is the slope of the type-2 zROC (= ratio of the unknown and known item evidence distribution standard deviations).
</aside>

The appendix of this page describes the R code used to implement this simulation.

### Example simulation run

```{r simulation-function}
sim_run <- function(p_know = NA, 
                    n_trials = 100, 
                    p_guess = 0,
                    d_gen = 1, 
                    s_gen = 1,
                    tau_gen = c(-Inf, -1, -.5, 0, .5, 1, Inf),
                    out = "pars",
                    ...) {

  levels <- length(tau_gen) - 1

  # Performance (p(know) to number of known items)
  if (is.na(p_know)) {
    # B & D: gaussian(50, 10) draw
    n_know <- round(rnorm(1, ceiling(n_trials/2), 10))
  } else if (p_know <= 1 & p_know >= 0) {
    # Or exact number known
    n_know <- ceiling(p_know * n_trials)
  }

  # Number of known items to number of accurate answers
  accuracy <- c(
    rbinom(n_trials - n_know, 1, p_guess),
    rep(1, times = n_know)
  )

  # Metacognition evidence values
  evidence <- c(
    rnorm(n_trials - n_know, 0, 1),
    rnorm(n_know, d_gen, s_gen)
  )

  # Evidence values to confidence ratings
  con6 <- cut(
    evidence, 
    breaks = tau_gen, 
    labels = FALSE
  )
  # Ratings to 0 - 5, as in data
  con6 <- con6-1

  # Return data or parameters
  dat <- data.frame(accuracy, con6)
  if (out == "pars") {
    metrics(dat, 
            confidence = NULL, 
            levels = levels,
            ...)
  } else if (out == "data") {
    dat
  }
}
```

Here is an example run of the simulation.

```{r sim-test, fig.cap="Example run of simulation scheme with $R = 1$: Histogram of simulated confidence ratings split by accuracy."}
tmp <- sim_run(
  out = "data", 
  d_gen = 1,
  p_know = .65,
  p_guess = 0.25,
  n_trials = nrow(dat)
)
metrics(tmp, confidence = NULL, ordinal=TRUE) %>% 
  kable(digits=2, caption = "Estimated metrics from example simulation run")
dat %>% 
  bind_rows(tmp, .id = "is_real") %>% 
  mutate(is_real = ifelse(is_real==1, "Real", "Simulated")) %>% 
  ggplot(aes(con6, fill = as.factor(accuracy))) +
  scale_fill_brewer("Accuracy", palette = "Set1") +
  scale_y_continuous("Count", 
                     expand = expand_scale(mult = c(0, .1))) +
  scale_x_continuous("Confidence", breaks = pretty_breaks(6)) +
  stat_count(aes(y=stat(count)),
             col="black", position = position_dodge()) +
  facet_wrap("is_real")
rm(tmp)
```

## Simulation

### Method

```{r simsp-generative-parameters, fig.cap = "Generative SDT model for simulation."}
Tau <-  c(-Inf, -1, -.5, 0, .5, 1, Inf)
sim_run(out="data", 
        d_gen=2, 
        tau_gen = Tau,
        n_trials=10000) %>% 
  table() %>% 
  kable(caption = "One set of generated data with d = 1")
p1 <- ggplot(data.frame(x=-3:6), aes(x)) +
  stat_function(fun = dnorm, size = 2) +
  stat_function(fun = dnorm, args = list(mean = .2), col = "red", size = 2) +
  geom_vline(xintercept = Tau, lty = 2) +
  theme(axis.title = element_blank())
p2 <- ggplot(data.frame(x=-3:6), aes(x)) +
  stat_function(fun = dnorm, size = 2) +
  stat_function(fun = dnorm, args = list(mean = 2), col = "red", size = 2) +
  geom_vline(xintercept = Tau, lty = 2) +
  theme(axis.title = element_blank())
(p1 + labs(title = "Poor metacognition (d = 0.2)")) / 
  (p2 + labs(title = "Good metacognition (d = 2)"))
```

We ran the simulation over a range of values of *K* and *R*. The simulation was repeated 100 times at each combination of the generative parameters, with 20000 trials for each repeat.

```{r sims-parameters-summary}
tibble(
  Variable = c(
    "Number of trials/items", 
    "Proportion of known items (knowledge)",
    "Guessing proportion (i.e. 1/m)",
    "Mean of evidence distribution for known items",
    "SD of evidence distribution for known items",
    "Evidence thresholds"
  ),
  Symbol = c(
    "$N_{trials}$",
    "K",
    "G",
    "R",
    "$\\sigma_k$",
    "$\\tau$"
  ),
  Values = c(
    20000, "0.05 - 0.95", "0 / 0.25", "0.05 - 2.45", 1.0, "-1, -.5, 0, .5, 1"
  )) %>% 
  kable(
    escape = FALSE, 
    caption = "Variable values used in main simulation run."
  )
```

```{r simsp-run}
# Performance and delta values to use in simulation
# hist(dat_metrics$da, breaks=100)
p_know_bins <- seq(0.05, .95, by = .1)
delta_bins <- seq(0.05, 2.45, by = .1)
# Define other parameters
sims <- as_tibble(
  expand.grid(
    rep = 1:100, 
    n_trials = 20000,
    d_gen = delta_bins,
    p_know = p_know_bins,
    p_guess = c(0, .25)
  )
)
if (!file.exists("data/sims.rda")) {
  plan(multiprocess)
  pars <- future_pmap_dfr(sims, sim_run, ordinal = FALSE, .progress = TRUE)
  beep(2)
  sims <- bind_cols(sims, pars)
  save(sims, file = "data/sims.rda")
} else { load("data/sims.rda") }

sims <- select(sims, d_gen:da)
sims_mean <- sims %>% 
  group_by(d_gen, p_know, p_guess) %>% 
  summarise(gamma = mean(gamma, na.rm = TRUE), da = mean(da, na.rm = TRUE)) %>% 
  ungroup()
sims_long <- sims_mean %>% 
  gather(parameter, estimate, gamma:da)
```

### Results

These simulations show that gamma and da are unrelated to performance when guessing=0. However, with guesses, both metrics are confounded by guessing: da and gamma are increasingly downward biased as performance decreases.

```{r main-figure, layout="l-body-outset", fig.width=7, fig.height=5.8}
sims_long %>% 
  # filter(d_gen %in% delta_bins[seq(1, length(delta_bins), 2)]) %>% 
  rename(Guessing = p_guess, Parameter = parameter) %>% 
  ggplot(aes(p_know, estimate, col=d_gen, group=d_gen)) +
  scale_color_viridis_c("Resolution") +
  geom_line(size=.75) +
  scale_x_continuous("Knowledge", breaks = pretty_breaks(5), limits = 0:1) +
  scale_y_continuous("Mean estimate", breaks = pretty_breaks(5)) +
  facet_grid(Parameter~Guessing, scales="free_y", labeller=label_both) +
  theme(legend.position = "bottom")
```

### Bias

We then assessed the amount of bias in the estimates (with guessing) at various levels of performance

```{r sims-bias-illustrate, layout="l-body-outset", fig.width=8, fig.height=6}
sims_long_bias <- sims_long %>% 
  spread(p_guess, estimate, sep = "") %>% 
  mutate()
sims_long_bias %>% 
  ggplot(aes(p_guess0, p_guess0.25, 
             col = p_know, group = p_know)) +
  scale_color_viridis_c("Performance") +
  geom_line(size = 1) +
  geom_abline(lty=2) +
  scale_x_continuous("Mean estimate (no guesses)", 
                     breaks = pretty_breaks(5)) +
  scale_y_continuous("Mean estimate (guesses)", 
                     breaks = pretty_breaks(5)) +
  theme(aspect.ratio=1, legend.position="bottom") +
  facet_wrap(~parameter, scales = "free", nrow=1)
```

## Compare data to simulations

We then compared the observed data values to the simulated values to examine to what extent the observed correlations resembled the necessary relations revealed by the simulation.

However, the simulated data were generated by varying underlying performance (proportion *known* items), and guesses were then added to produce the observed accuracy values. Accuracy values in the data, likewise, are contaminated by guesses. One way to correct for this is to use a formula to try to convert the observed proportion correct to underlying performance. Thus, we can "correct" the guess-contaminated observed proportions correct to performance values using e.g. Abbott's formula.

<aside>
Abbott's formula:
$$
\hat{p}_c = \frac{\hat{p} - .25}{1 - .25}
$$
</aside>

```{r guessing-correct-plot, fig.cap = "Comparison of observed and corrected proportion corrected."}
# Illustrate guessing correction
dat_metrics %>% 
  ggplot(aes(p, p_know)) +
  geom_point(shape=1) +
  geom_abline(lty = 2) +
  coord_cartesian(ylim = 0:1, xlim = 0:1) +
  labs(x="Observed proportion correct", y="Corrected p(correct) ('knowledge')") +
  theme(aspect.ratio=1)
```

### Regression test

We then tested whether the performance-resolution metric regression slopes were significantly steeper in the data, vs the equivalent slope in the simulations.

We tested

$$
\frac{ \hat{\beta}_{1, data} - \hat{\beta}_{1, simulation} }
{s.e.\hat{\beta}_{1, data}}
$$

```{r regression-test-fig, layout="l-page", fig.width=8, fig.height=3.6, fig.cap="Comparison of observed correlations to simulation results."}
DEGREE <- 1
# PLot
tmp <- dat_metrics %>% 
  group_by(exp) %>% 
  summarise(d_ord = mean(d_ord, na.rm=T)) %>% 
  group_by(exp) %>% 
  mutate(d_gen = delta_bins[which.min(abs(d_ord - delta_bins))]) %>% 
  inner_join(sims_long) %>% 
  filter(p_guess == 0.25)
tmp %>% 
  ggplot(aes(p_know, estimate)) +
  scale_x_continuous(
    "Knowledge", 
    breaks = pretty_breaks(5)
  ) +
  scale_y_continuous(
    "Estimate", 
    breaks = pretty_breaks(5)
  ) +
  theme(legend.position = "none") +
  geom_point(
    data = gather(mutate(dat_metrics, gamma=gz), parameter, value, da, gamma), 
    aes(x = p_know, y = value), 
    col = "black", fill = "white", shape = 21, size = .5, alpha = .8
  ) +
  geom_smooth(
    data = gather(mutate(dat_metrics, gamma=gz), parameter, value, da, gamma), 
    aes(x = p_know, y = value), lty=2, 
    formula = y ~ poly(x, DEGREE),
    method = "lm", size=.75, se=T, alpha=.33, col="black"
  ) +
  geom_smooth(
    method = "lm", aes(col=d_gen), se=F, 
    formula = y ~ poly(x, DEGREE)
  ) +
  scale_color_viridis_c(
    "Resolution", 
    values = rescale(tmp$d_gen, to = c(0, 1))
  ) +
  facet_grid(parameter~exp, scales="free_y") +
  theme(panel.spacing = unit(11, "pt"))
```

And then the tests

```{r regression-test}
# Simulated slope at each experiment's mean d_ord
tmp1 <- dat_metrics %>% 
  group_by(exp) %>% 
  summarise(d_ord = mean(d_ord, na.rm=T), s = mean(s, na.rm=T)) %>% 
  group_by(exp) %>% 
  mutate(d_gen = delta_bins[which.min(abs(d_ord - delta_bins))]) %>% 
  inner_join(sims_long_bias) %>% 
  ungroup() %>% 
  nest(-exp, -parameter) %>% 
  mutate(
    res = map(data, ~lm(p_guess0.25 ~ p_know, data = .)),
    out = map(res, ~tidy(.))
  ) %>% 
  unnest(out) %>% 
  select(exp, parameter, term, sim_coef = estimate)

# Observed slope for each experiment
tmp2 <- dat_metrics %>% 
  mutate(gamma = gz) %>% 
  gather(parameter, value, gamma:da) %>% 
  nest(-exp, -parameter) %>% 
  mutate(
    res = map(data, ~lm(value ~ p_know, data = .)),
    out = map(res, ~tidy(.)),
    N_subj = map_dbl(data, ~nrow(.))) %>% 
  unnest(out, .drop = T) %>%
  select(exp, N_subj, parameter, term, data_coef = estimate, std.error)

# Test each experiment's slope against sim slopes
left_join(tmp1, tmp2) %>% 
  mutate(
    tval = (data_coef - sim_coef) / std.error,
    # One-sided p-values
    pval = pt(tval, df = N_subj-2, lower=FALSE)
  ) %>% 
  filter(!(term %in% c("p_know2", "(Intercept)"))) %>%
  arrange(parameter) %>% 
  kable(
    digits = 2, 
    caption = "Tests of data regression slopes against simulations."
  )
```

### Test subjects against simulated curve at perfect metacognition

Here, we attempt to test observed gamma/da against near-perfect metacognition. Each participant's observed gamma/da is paired with the corresponding mean simulated estimate at maximum d_gen (corresponding to about gamma = .96) and guessing = 0.25 at the participant's level of performance.

```{r}
tmp1 <- dat_metrics %>% 
  select(exp, subject, gamma = gz, da, p_know) %>% 
  group_by(exp, subject) %>% 
  mutate(p_know_bin = p_know_bins[which.min(abs(p_know - p_know_bins))]) %>% 
  ungroup()
tmp2 <- sims_mean %>% 
  filter(d_gen == max(delta_bins), p_guess == "0.25")
tmp <- left_join(tmp1, tmp2, by = c("p_know_bin" = "p_know"))
names(tmp) <- str_replace_all(names(tmp), "\\.x", ".data")
names(tmp) <- str_replace_all(names(tmp), "\\.y", ".perfect")
mean(tmp$gamma.data)
mean(tmp$gamma.perfect)
# Are subjects gamma/da significantly lower than the gamma/da from simulation with perfect metacognition? That is, we use the difference between each subject's gamma and the simulated perfect gamma at that subjects point of performance
t.test(tmp$gamma.perfect, tmp$gamma.data, paired = T)
t.test(tmp$da.perfect, tmp$da.data, paired = T)
p1 <- tmp %>% 
  mutate(
    gamma_diff = gamma.perfect - gamma.data,
    da_diff = da.perfect - da.data
  ) %>% 
  select(exp, subject, contains("_diff")) %>% 
  gather(key, value, gamma_diff, da_diff) %>% 
  ggplot(aes(value)) +
  geom_histogram(bins = 50) +
  facet_wrap("key", scales = "free")
p2 <- tmp %>% 
  ggplot(aes(gamma.perfect, gamma.data)) +
  geom_abline(lty = 2) +
  geom_point()
p3 <- tmp %>% 
  ggplot(aes(da.perfect, da.data)) +
  geom_abline(lty = 2) +
  geom_point()
(p3 | p2) / p1
```

```{r}
# Difference from perfect on performance
# (How much below perfect are they as function of performance)
p4 <- tmp %>% 
  mutate(
    gamma_diff = gamma.perfect - gamma.data,
    da_diff = da.perfect - da.data
  ) %>% 
  select(exp, subject, p_know, contains("_diff")) %>% 
  gather(key, value, gamma_diff, da_diff)
p4 %>% 
  ggplot(aes(p_know, value)) +
  geom_point() +
  geom_smooth() +
  facet_wrap("key", scales = "free_y")
```

## Appendix A: Replication of Benjamin and Diaz

We replicated the simulation of @benjamin2008measurement: For several levels of generative *R*, we ran 100 sim-subjects with 2000 trials each. For each subject, *K* was drawn from $N(50, 10^2)$, rounded to an integer. Instead of 4 rating categories, we used 6 to more closely correspond to our data. *G* was varied at 0 and 0.25. $\sigma_k$ was either 1 or 1.5.

```{r simsbd-run}
sims_bd <- as_tibble(
  expand.grid(
    sub = 1:100, 
    n_trials = 2000,
    d_gen = seq(0, 2.5, length.out = 10),
    s_gen = c(1, 1.5),
    p_guess = c(0, 0.25)
  )
)

# Run & save / load simulation from disk
if (!file.exists("data/sim-B&D.rda")) {
  plan(multiprocess)
  pars <- future_pmap_dfr(sims_bd, sim_run, ordinal = TRUE, .progress = TRUE)
  beep(2)
  sims_bd <- bind_cols(sims_bd, pars)
  save(sims_bd, file = "data/sim-B&D.rda")
} else { load("data/sim-B&D.rda") }
```

```{r simsbd-plot1, layout="l-body-outset", fig.width=8, fig.height=6, fig.cap="Results from simulation replicating Benjamin and Diaz's (2008) simulation scheme when guessing=0. The two columns indicate different generative $\\sigma_s$s."}
sims_bd %>% 
  filter(p_guess == 0) %>% 
  gather(parameter, value, gamma, da, y0, da_ord, d_ord, r) %>% 
  group_by(parameter, d_gen, s_gen) %>% 
  summarise(n = n(), mean = mean(value, na.rm = T)) %>% 
  ggplot(aes(d_gen, mean, col = parameter)) +
  scale_color_brewer(
    "Parameter", 
    palette = "Dark2", 
    aesthetics = c("color", "fill")
    ) +
  geom_line(size=1.25) +
  geom_abline(lty=2, size = .33) +
  scale_x_continuous("Resolution", breaks = pretty_breaks(5)) +
  scale_y_continuous("Mean estimate", breaks = pretty_breaks(5)) +
  facet_wrap("s_gen", labeller=label_both) +
  theme(aspect.ratio = 1)
```

The results of this simulation mirrored those of @benjamin2008measurement: Both gamma and $d_a$ are monotonically increasing with generative *R*, but only $d_a$ has the property of scaling linearly with it. Second, the SDT metrics are more accurately estimated via an ordinal regression [@KnoblauchModelingPsychophysicalData2012]. The $d'_{ord}$ metric recovers the generative parameter in every simulation, and its transformations ($d_{a-ord}$) are easily obtained and mirror those of the line-fitting procedure.

### Introducing guessing

However, when some correct answers are guesses (Figure \@ref(fig:simsbd-plot-guessing), below), all resolution metrics considered here are biased downward.

```{r simsbd-plot-guessing, layout="l-body-outset", fig.width=8, fig.height=6, fig.cap="Results from simulation replicating Benjamin and Diaz's (2008) simulation scheme when guessing=0.25. The two columns indicate different generative $\\sigma_s$s."}
sims_bd %>% 
  filter(p_guess == 0.25) %>% 
  gather(parameter, value, gamma, da, y0, da_ord, d_ord, r) %>% 
  group_by(parameter, d_gen, s_gen) %>% 
  summarise(n = n(), mean = mean(value, na.rm = T)) %>% 
  ggplot(aes(d_gen, mean, col = parameter)) +
  scale_color_brewer(
    "Parameter", 
    palette = "Dark2", 
    aesthetics = c("color", "fill")
    ) +
  geom_line(size=1.25) +
  geom_abline(lty=2, size = .33) +
  scale_x_continuous("Resolution", breaks = pretty_breaks(5)) +
  scale_y_continuous("Mean estimate", breaks = pretty_breaks(5)) +
  facet_wrap("s_gen", labeller=label_both) +
  theme(aspect.ratio = 1)
```

Because this pattern of bias is general to all metrics considered here, for simplicity we only consider gamma and $d_a$ below, because they have most often been discussed in previous literature.

For completeness, we compared estimated $\hat{d_a}$ to generative $d_a$

```{r}
sims_bd %>% 
  mutate(da_gen = (sqrt(2) * (d_gen/s_gen)) / sqrt(1 + (1/s_gen)^2)) %>% 
  ggplot(aes(da_gen, da, 
             col = as.factor(p_guess), 
             fill = as.factor(p_guess))) +
  geom_abline(lty=2) +
  stat_summary(geom="ribbon") +
  facet_wrap("s_gen", labeller=label_both) +
  theme(aspect.ratio=1)
```

## Appendix B: Simulation function {.appendix}

```{r show-simulation-function, ref.label="simulation-function", echo=TRUE}
```

## Appendix C: Calculating gamma from ROC {.appendix}

```{r, layout="l-body-outset"}
# Gamma can be calculated from AUC (Higham & Higham, 2018)
tmp <- sims_bd %>% 
  filter(p_guess == 0) %>% 
  select(-sub, -n_trials, -r, -p, -n, -con_m) %>% 
  mutate(gamma_true = 2 * pnorm((d_gen / s_gen) / sqrt(1 + (1/s_gen)^2)) - 1) %>% 
  mutate(az_zroc = pnorm(y0 / sqrt(1 + (1/s)^2 )),
         gamma_zroc = 2*az_zroc-1) %>%
  mutate(az_ord = pnorm( (d_ord/s_ord) / sqrt(1+ (1/s)^2 )),
         gamma_ord = 2*az_ord-1)

tmpp <- tmp %>%   
  filter(s_gen == "1") %>% 
  ggplot(aes(gamma_true, col = d_gen)) +
  scale_colour_viridis_c() +
  geom_abline(lty=2) +
  stat_summary(fun.y=mean, geom = "line") +
  geom_point() +
  theme(aspect.ratio = 1, legend.position = "bottom")
(tmpp + aes(y=gamma)) /
  ((tmpp + aes(y=gamma_zroc) | 
      tmpp + aes(y=gamma_ord)) & 
     theme(legend.position="none"))

tmpp2 <- dat_metrics %>% 
  select(-r, -p, -con_m) %>% 
  mutate(az_zroc = pnorm(y0 / sqrt(1 + (1/s)^2 )),
         gamma_zroc = 2*az_zroc-1) %>%
  mutate(az_ord = pnorm( d_ord / sqrt(1 + (1/s_ord)^2 )),
         gamma_ord = 2*az_ord-1) %>% 
  ggplot(aes(gamma)) +
  geom_abline(lty=2) +
  theme(aspect.ratio=1)
(tmpp2 + geom_point(aes(y=gamma_zroc)) | 
  tmpp2 + geom_point(aes(y=gamma_ord))) &
  coord_cartesian(ylim=-1:1, xlim = -1:1)

```